### 2022-12-17 Rmd/paper 02 Scale-jump analysis of historical retrospective patterns.Rmd
Restructured into Rmd folder

# -----------------------------------------------------------------------------------------------
# Analyse changes in pelagic stocks.Rmd
#
# 07/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
# 13/01/2019 now only figures and tables
# 15/01/2019 added scale jump figure with explanatory variables
# 22/02/2019 focussed on Pairwise Comparison of Means = scale jumps under different conditions
# -----------------------------------------------------------------------------------------------

@@

mynyears      <- 5
myremoveyears <- 0
myvar         <- "stocksize"
mytitle       <- paste0(toupper(myspecies), ", ", myvar, ", removed years: ", myremoveyears)

@@

*Figure 6: Comparisons of stock size (SSB) in assessments where stock definitions have changed. Blue-green is old stock definition, red is new stock definition*

@@

@@

*Figure 7: Number of scale jumps in stocksize for the nine stocks considered, using 5 year means and a threshold of 10%. The red line indicates the number of assessments included in the comparison (right hand y-axis)*

@@

*Figure 8: Proportion of scale jumps in stocksize by stock and decade, using 5 year means and a threshold of 10%. The red line indicates the number of assessments included in the comparison (right hand y-axis)*

@@

# scaling of the two axes
ylim.prim <- c(0, 100)   # number of assessments  
ylim.sec  <- c(0, 1)     # proportions

>>>>

# -----------------------------------------------------------------------------------------------
# Paper 02 Scale-jump analysis of historic retrospective patterns.Rmd
#
# 07/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
# 13/01/2019 now only figures and tables
# 15/01/2019 added scale jump figure with explanatory variables
# 22/02/2019 focussed on Pairwise Comparison of Means = scale jumps under different conditions
# -----------------------------------------------------------------------------------------------

@@

mynyears      <- 5
myremoveyears <- 0
myvar         <- "stocksize"
# mytitle       <- paste0(toupper(myspecies), ", ", myvar, ", removed years: ", myremoveyears)

@@

*Figure 7: Comparisons of stock size (SSB) in assessments where stock definitions have changed. Blue-green is old stock definition, red is new stock definition*

@@

# *Figure 7: Number of scale jumps in stocksize for the nine stocks considered, using 5 year means and a threshold of 10%. The red line indicates the number of assessments included in the comparison (right hand y-axis)*

@@

```{r eval=FALSE, fig.align="center", fig.asp=0.8, message=FALSE, warning=FALSE, include=FALSE}

@@

```{r echo=FALSE, fig.align="center", fig.asp=0.4, message=FALSE, warning=FALSE}

# plot number of scale jumps by decade

t <-
  res$d.pcm %>% 
  filter(variable        == myvar,
         numberofyears   == mynyears,
         removelastyears == myremoveyears,
         threshold       == mythreshold) %>% 
  distinct(stock, variable, assessmentyear, phi, scale_jump, 
           removelastyears, numberofyears, threshold) %>%
  filter(!is.na(scale_jump)) %>% 
  mutate(decade = (10*floor(assessmentyear/10)) ) %>% 
  # mutate(decade = paste0(decade, "s")) %>% 
  group_by(variable, decade, scale_jump) %>% 
  summarize(n = n()) %>% 
  group_by(variable, decade) %>% 
  mutate(prop = n / sum(n), 
         nassess = sum(n)) %>% 
  filter(scale_jump == TRUE)

p1 <-
  ggplot(t, aes(x=reorder(decade, -decade), y=prop)) +
  theme_publication() +
  theme(
    panel.spacing    = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), 
                                    angle = 180),
    # plot.background   = element_rect(fill="red"),
    plot.margin        = unit(c(0,0,0,0.0),"cm"),
    strip.placement = "outside",
    legend.position  = "none"
  ) +

  geom_bar(stat="identity") +
  scale_y_continuous("Proportion scale jumps", breaks=pretty_breaks(n=4)) +
  labs(x="") +
  coord_flip()

p2 <-
  t %>% 
  ggplot(aes(x=reorder(decade, -decade), y=nassess)) +
  theme_publication() +
  theme(
    panel.spacing      = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background   = element_blank(),
    strip.text.x       = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
    strip.text.y       = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), 
                                    angle = 180),
    axis.text.y = element_blank(), 
    axis.ticks.y=element_blank(),
    plot.margin        = unit(c(0,0,0,0.0),"cm"),
    # plot.background   = element_rect(fill="blue"),
    strip.placement    = "outside",
    legend.position    = "none"
  ) +

  geom_path(colour="red") +
  geom_point(colour="red") +
  expand_limits(y=0) +
  labs(x="") +
  coord_flip()

cowplot::plot_grid(p1, 
                   p2 , 
                   ncol=2,
                   scale=0.99, align="hv", 
                   rel_widths = c(1.2,0.6))

# scaling of the two axes
# ylim.prim <- c(0, 0.5)     # proportions
# ylim.sec  <- c(0, 90)     # number of assessments

# calculations for two axes;
# https://stackoverflow.com/questions/3099219/ggplot-with-2-y-axes-on-each-side-and-different-scales
# b <- diff(ylim.prim)/diff(ylim.sec)
# a <- b*(ylim.prim[1] - ylim.sec[1])

# t %>%   
#   ggplot(aes(x=reorder(decade, -decade), y=prop)) +
#   theme_publication() +
#   theme(
#     panel.spacing    = unit(0.1, "lines"),
#     panel.grid.major.x = element_line(colour = "gray80"),
#     strip.background = element_blank(),
#     strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
#     strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), 
#                                     angle = 180),
#     strip.placement = "outside",
#     legend.position  = "none"
#   ) +
#   
#   geom_bar(stat="identity") +
#   geom_line(data=t2, aes(x= reorder(decade, -decade), y = a + nassess * b), color = "red", size=1) +
#   scale_y_continuous("Proportion scale jumps", 
#                      sec.axis = sec_axis(~ (. - a)/b, 
#                                          name = "N Assessments", 
#                                          breaks=pretty_breaks(n=4))) +
#   labs(x="") +
#   coord_flip()
#

*Figure 6: Proportion of scale jumps in stocksize by decade, using 5 year means and a threshold of 10%. The red line indicates the number of assessments included in the comparison (right hand y-axis)*

@@

p1 <-
  t %>%

  ggplot(aes(x=reorder(assessmentmodel, -rank), y=prop)) +
  theme_publication() +
  theme(
    panel.spacing    = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), 
                                    angle = 180),
    # axis.text.y = element_blank(), 
    # axis.ticks.y=element_blank(),
    plot.margin        = unit(c(0,0,0,0.0),"cm"),
    strip.placement = "outside",
    legend.position  = "none"
  ) +

  geom_bar(stat="identity") +
  labs(x="") +
  scale_y_continuous("Proportion scale jumps", breaks=pretty_breaks(n=4)) +
  coord_flip()

p2 <-
  t %>%

  ggplot(aes(x=reorder(assessmentmodel, -rank), y=nassess)) +
  theme_publication() +
  theme(
    panel.spacing    = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm")),
    strip.text.y     = element_text(face="bold", hjust=0, margin = margin(2,0,2,0, "mm"), 
                                    angle = 180),
    axis.text.y = element_blank(), 
    axis.ticks.y=element_blank(),
    plot.margin        = unit(c(0,0,0,0.0),"cm"),
    strip.placement = "outside",
    legend.position  = "none"
  ) +

  geom_point(colour="red") +
  geom_line(colour="red") +
  labs(x="") +
  scale_y_continuous("N assessments", breaks=pretty_breaks(n=4)) +
  coord_flip()

cowplot::plot_grid(p1, 
                   p2 , 
                   ncol=2,
                   scale=0.99, align="hv", 
                   rel_widths = c(1.2,0.6))

# scaling of the two axes
# ylim.prim <- c(0, 100)   # number of assessments  
# ylim.sec  <- c(0, 1)     # proportions

# calculations for two axes;
# https://stackoverflow.com/questions/3099219/ggplot-with-2-y-axes-on-each-side-and-different-scales
# b <- diff(ylim.prim)/diff(ylim.sec)
# a <- b*(ylim.prim[1] - ylim.sec[1])
